Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-04-23 manuscript/manuscript.tex

First, let us review how the KS statistic is typically computed for a sample
with no serial dependence with fitted parameters.
Let $\hat\theta_n$ denote the fitted parameters for the hypothesized 
distribution $F(x; \hat\theta_n)$ fitted onto $X_t$, and let 
$F_n$ denote the empirical distribution function based on $X_1,...,X_n$.
Let
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}

Then, the
goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

\jy{State clearly the test statistic definition. Point out that the null
  distribution depends on the unknown parameters and unknown serial
  dependence. Point out the challenges are to approximate the null distribution.}
\mc{addressed in third paragraph of introduction}

We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
\jy{clean up the notations. Has $F$ been defined yet? do you need cdf?}
\mc{addressed}
where $F(x)$ is the true cumulative distribution function 
(under the $H_0: F(x) = F(x; \theta_0)$ for some true
parameter $\theta_0$.)

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
\jy{define $\Phi$}\mc{addressed}
\jy{The language issues are more than I expected. Please do a careful proofread,
  now that you are more experienced.}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution and $\Phi$ is the distribution function of the standard normal
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 
\break 0.639, 0.861\}$.

@@

Using the \textsl{qqplotr} and \textsl{ggplot2} packages 
\citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, the
points will be aligned with the diagonal line. This 
indicates that under the null hypothesis, our method will only reject the $H_0$
at the rate of the significance level.

\jy{Remove the white space in the pdf figures. Could put the zoomed version in
  the same figure.}\mc{addressed}

@@

Under $H_0$, the p-values for the non-parametric block bootstrap KS test are
\jy{Not true from the plots.}\mc{addressed}
uniformly distributed as long as the serial dependence is not too high. 
This is an indication that our method 
holds its size for moderate serial dependence levels.
For such cases, we can expect the probability
of a type 1 error to be approximately equivalent to the significance level.

@@

We have also 
demonstrated possible applications of this method to financial data.
\jy{don't astrophie citet; let alone citep}\mc{addressed}
Through comparison with the non-parametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric
bootstrap method developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}\mc{addressed}
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA processes. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

>>>>

First, let us review how the KS statistic is computed for an independent
sample with fitted parameters. Let $\hat\theta_n$ denote the parametrically
fitted parameters, which can be obtained from any consistent estimator with
asymptotic normality properties,  Let $F_n$ be the empirical distribution
function based on $X_1,...,X_n$.
Define 
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the one-sample KS goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

Because of the estimation uncertainty in $\hat\theta_n$, the asymptotic
distribution of $Y_n$ is no longer distribution-free. This is in contrast to the
case where $\theta$ is known.  Note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)).
\end{equation*}
When $\hat\theta_n$ is replaced with $\theta$, as in the standard KS testing
situation, the second term vanishes, and $T_n$ converges in distribution to that
of $\sup_t | B(F(t)) |$, where $B(\cdot)$ is the standard Brownian bridge
\citep{kolmogorov1933sulla}. With $\hat\theta_n$ in place of $\theta$, the
second term leads to a bias in the Brownian bridge, which was addressed by
\citet{babu2004goodness}.

When $\{X_t: t = 1, \ldots, n\}$ are further serially dependent, the statistics
$T_n$ still provides a reasonable metric for deviation from the null
hypothesis. Nonetheless, the asymptotic distribution of $T_n$ cannot be fixed
even with the bias correction of \citet{babu2004goodness}, because the
correction requires independent sample.

>>>>>>> 8ae0002af5827811aa9091e769fc16303007bc1c

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
\jy{The language issues are more than I expected. Please do a careful proofread,
  now that you are more experienced.}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution and $\Phi$ is the distribution function of the standard normal
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are
\[
  \{-0.884, -0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}.
\]

@@

Using R packages \textsl{qqplotr} and \textsl{ggplot2} \citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, the
points will be aligned with the diagonal line. This 
indicates that under the null hypothesis, our method will only reject the $H_0$
at the rate of the significance level.
\jy{Describe the table not zoomed-in figure.}
A zoomed in plot for probabilities between 0 and 0.1 is
also provided, as
that is the most common range for significance levels $\alpha$.

@@

Under $H_0$, the p-values for the non-parametric block bootstrap KS test are
\jy{The flow doesn't read right from the last paragraph}
uniformly distributed as long as the serial dependence is not too high. 
This is an indication that our method 
holds its size for moderate serial dependence levels.
For such cases, we can expect the probability
of a type 1 error to be approximately equivalent to the significance level.

@@

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with the non-parametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric
bootstrap method developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}\mc{addressed}
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA processes. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.
